---
title: "Homework 5 - Unsupervised"
author: '[Your name here]'
date: "Due 7 December at 11pm"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = TRUE)
library(tidyverse)
library(testthat)
library(Stat406)
theme_set(theme_bw())
```

### Instructions {.unnumbered}

* Use the space inside of 

```
::: {.solbox data-latex=""}


:::
```

to answer the following questions.


* Do not move this file or copy it and work elsewhere. Work in the same
directory. 
* Use a new branch named whatever you want. Create it now! Can't come up with 
something, try [here](https://alator21.github.io/repository-name-generator/).  
Make a small change, say by adding your name. Commit and push now!
* Try to Knit your file now. Are there errors? Fix them now, not at 11pm on the 
due date.
* There MUST be some text between `::: {.solbox}` and the next `:::` or this 
will fail to Knit. 
* If your code or figures run off the edge of the `.pdf`, you'll lose 2 points 
automatically.
* Be sure to add your name in the `author` field at the top.


\clearpage


# Introduction {-}

Throughout this assignment, we will be looking at a data set of Whisky flavour 
profiles. David Wishart was at one point, Chief Statistician at the Scottish 
Office of the Civil Service. When he retired, he focused on Whisky and wrote a 
book called ["Whisky Classified"](https://www.powells.com/book/whisky-classified-choosing-single-malts-by-flavour-9781911595731). 
For the book, he collected tasting notes published about 86 different Scotch 
Whisky distilleries on a number of aspects and "distilled" them down to 12 
flavour categories. Then each distillery's representative whisky was given a 
score on each category from 0-4, 0 meaning that that flavour is not represented 
in that whisky, 4 meaning that it is strongly represented. 

The data set was later expanded to include more distilleries and crowd-sourced 
tasting notes, but, this data seems to be kept only in a for-profit Windows 
software which no longer exists. Dr. Wishart passed away in 2020, and there 
seems to be no way to access the larger data set.

You have here, the version of the data from the first edition of Dr. Wishart's 
book. An article describing some of his analyses with 185 single malts was 
published in 2009 in [Significance](https://rss.onlinelibrary.wiley.com/doi/full/10.1111/j.1740-9713.2009.00337.x). 

We will undertake some similar analyses in this homework assignment. A snapshot 
of the data is shown below.

```{r whisky-vars, echo=FALSE}
N <- nrow(whisky)
whisky %>% str()
```


\clearpage

# Dimension reduction (6 points)

## (0.5 points) 

Create a bar chart that shows 12 panels, one for each of the twelve flavour
profiles. In each panel, the heights of the bars should be the percent of 
whiskys with each score 0-4 (out of 86 total) (the y-axis is a percent between 
0 and 100). Make sure that your graphic fits on 1 page.

::: {.solbox data-latex=""}

```{r flavorings}
# some code.
```


:::



## (0.5 points) 

Charts should always be described in any analysis or else they are useless. 
Examining your chart, describe any patterns you see in a few sentences 
(at most 3). There are many right answers. Wrong answers are things like 
"there are 12 panels" or "each panel has 5 bars" or "I don't see anything".


::: {.solbox data-latex=""}

Some text.

:::



## (1 point) 

Perform PCA on the 12 flavour categories. Produce a scree plot with the y-axis
displaying the percent of total variance explained by each component. Based on 
your plot, is 2 a reasonable number of dimensions to use? Why or why not?

::: {.solbox data-latex=""}
```{r estimate-pca}
# some code

```

Some text.

:::



## (1 point)

Produce a plot of PC1 vs PC2. Instead of a dot for each distillery, show the 
name of the distillery in the figure. (You'll have lots of words, some of which 
may be hard to read). Here's example code to get the idea. 

```{r, echo=TRUE, eval=FALSE}
plot(1:5, 1:5, pch = "")
text(1:5, 1:5, labels = letters[1:5])
```

In `ggplot`, you would use `geom_text()` or `geom_label()` rather than
`geom_point()`. 
    
Also produce a plot of the Loadings for the first two PCs. Which qualities seem
important (positive or negative) for each PC (describe this in 1-2 sentences)? 
Do any Distilleries stick out, if so, what qualities do they have in common?

::: {.solbox data-latex=""}


```{r, pc1-v-pc2}
# Code to plot pc1 against pc2

```

```{r weights-plot}

# code to plot the weights

```

Some text.
:::




## (2 points) 

Complete the function below. See the documentation for help with the arguments.
Your function must return a
matrix of dimension `n` x `M` where `n` is the number of rows of `K`. Do not 
change the function signature.   

You should make your function produce errors or warnings if invalid inputs are
passed. This is easiest way to do with `stop()` or `stopifnot()` or `warning()`.
Try examining the documentation for those functions. Looking at the tests should
help you determine what sorts of checks to perform and what warnings to throw.
Set `eval=TRUE` in the chunk options when your function is ready.

**Hint: think carefully about what the sign of `tol` should be to make this**
**work. It may not conform to your first instinct.**

::: {.solbox data-latex=""}
```{r kpca-function}
#' @param K a symmetric, non-negative definite Kernel matrix
#' @param M positive integer for the target embedding dimension 
#' @param tol tolerance to test for negative definiteness.
kpca <- function(K, M = 2, tol = -1e-8) {
  
  
  
  
}
```

```{r kpca-tests, eval=FALSE, echo=FALSE}
# Don't change this chunk. Set eval=TRUE when your function is ready.
z <- 1:5
zz <- matrix(c(1:5, 2, -1, 3, -2, 4), ncol = 2)
p <- matrix(-.2, 5, 5) + diag(5)
e <- eigen(p %*% zz %*% t(zz) %*% p)
ee <- e$vectors[,1:2] %*% diag(sqrt(e$values[1:2]), 2, 2)
test_that("Kernel PCA is reasonable", {
  expect_error(kpca(whisky))
  expect_warning(kpca(z %o% z))
  expect_error(kpca(z))
  expect_error(kpca(matrix(1:25, 5, 5)))
  expect_error(kpca(matrix(1:24, 4, 6)))
  expect_warning(kpca(zz %*% t(zz), .9))
  expect_warning(o <- kpca(diag(2), 2))
  expect_equal(o, sqrt(2) / 2 * matrix(c(-1,1,0,0), 2))
  expect_warning(o <- kpca(matrix(c(1,-1,-1,1), 2)))
  expect_equal(o, matrix(c(-1,1,0,0), 2))
  expect_equal(kpca(zz %*% t(zz)), ee)
})
```
:::





## (1 point) 

Use your `kpca()` function to estimate kernel PCA. Form your `K` by first
calculating the distance matrix between rows in your data using the `canberra`
distance. The easiest way is with the `dist()` function. You have to convert the
result to a matrix with `as.matrix()`. Then set `K` to be `1 -` this result. 
Pass that in to your function. Use `M = 2`.   

Now the hard part, produce a 12 panel plot as in 1. The x-axis should be the 
first component, the y-axis should be the second component. Colour the points 
based on the flavour score (0-4). Describe any patterns you notice in 2-3
sentences.

::: {.solbox data-latex=""}
```{r whisky-kpca}
# Some code.

```

Some text.

:::

\clearpage

# Clustering (4 points)

## (1 point) 

Use `kmeans()` to cluster whiskys using the 12 flavour metrics. Try K from 2 to 
20. Produce a plot of the CH Index against K. 


::: {.solbox data-latex=""}

```{r kmeans-ch}
set.seed(406406406) # don't change me
K <- 2:20



```

:::


## (.5 points) 

Dr. Wishart used 10 clusters to describe these whiskys. Does this seem like a
justifiable choice given the CH index? Why or why not?

::: {.solbox data-latex=""}

Some text.

:::

## (1 point) 

Make a plot of PC1 vs PC2 as in Question 1.5 above, but colour the distillery 
names by cluster assignment. (You should be able to copy the code from above and make some minor changes.) Use the number of clusters that maximizes the CH Index. Describe any patterns you see.

::: {.solbox data-latex=""}
```{r cluster-pcs}
# Some code.

```

Some text.

:::



## (0.5 points) 

Use the function below to create a map of Scotland with the distilleries plotted
as points coloured by their cluster assignment (the argument `cl` should be a
vector of cluster assignments). Do this for the best K (as determined by CH 
index) and for K = 10. Describe the patterns you see. 

```{r scotland-plot-fun}
scotland_cluster_plot <- function(cl) {
  stopifnot(is.vector(cl), length(cl) == N)
  ggplot(whisky %>% mutate(cluster = cl), aes(Longitude, Latitude)) +
    geom_point(aes(colour = factor(cluster))) + 
    coord_sf(ylim = c(53, 60), xlim = c(-7, -1)) + 
    borders(regions = "UK") +
    scale_colour_viridis_d() +
    theme(legend.position = "none")
}
```

::: {.solbox data-latex=""}
```{r}
# Some code.

```

some text

:::



## (1 point) 

Ron Swanson (of [Parks and Recreation](https://www.imdb.com/title/tt1266020/), 
see the image below) famously loves Lagavulin (as do I). In light of your 
analyses in this assignment, are there other whiskys you would recommend Ron 
try? What flavour profiles does Ron enjoy? Produce any figures that can be used
to back up your claims.

```{r ron, fig.align='center'}
knitr::include_graphics("ron.pdf")
```


::: {.solbox data-latex=""}

Some text.

```{r}

# Code as needed.

```



:::
